Spatial phylogenetics of Japanese ferns: Patterns, processes, and implications for conservation

### 2021-09-21 ms/manuscript.Rmd
WIP convert to AJB format

---
title: "Spatial phylogenetics of Japanese ferns: Patterns, processes, and implications for conservation"
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::word_document2:
    pandoc_args: [ "--csl", "apa-6th-edition.csl"]
    toc: no
    number_sections: no
    reference_docx: template.docx
  bookdown::pdf_document2:
    citation_package: default
    toc: no
    number_sections: no
    fig_caption: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "apa-6th-edition.csl"]
params:
  doc_type: doc
editor_options: 
  chunk_output_type: console
---

@@

Joel H. Nitta^1^, Brent D. Mishler^2^, Wataru Iwasaki^1,3^, and Atsushi Ebihara^4^

@@

Author for correspondence:  
*Joel H. Nitta*  
*Email: joelnitta@gmail.com*

@@

**Aim:** To characterize multiple facets and drivers of biodiversity and understand how these relate to bioregions and conservation status.

**Location:** Japan

**Taxon:** Ferns

@@

**Main conclusions:** The integrative approach used here reveals previously undetected patterns and drivers of biodiversity in the ferns of Japan. 
The different distributions of each diversity type reflect the major bioregions. 
Future conservation efforts should recognize that threats can vary by biodiversity metric and consider multiple multiple metrics when establishing conservation priorities.

@@

## Introduction

@@

## Materials and Methods

@@

## Results

@@

Taxonomic bioregions 3 and 4 have higher SES values of PD, RPD, FD, and RFD than bioregions 1 and 2 (`r figure("div-by-tax-region")`). 
Bioregions 3 and 4 also have consistently high PE *p*-scores, whereas bioregions 1 and 2 have a much wider variation. 
Bioregion 2 has lower RPD and RFD than bioregion 1. 
Comparison of biodiversity metrics across phylogenetic bioregions was similar, except that phylogenetic bioregion 3 mostly corresponds to taxonomic bioregions 3 and 4 combined.

@@

## Discussion

@@

### Limits of the study and future outlook

@@

There are a number of caveats that must be kept in mind when interpreting these results, but these also suggest avenues for future research. 
First, the data include specimens collected over a wide time period, so they do not necessarily represent the current distribution of ferns in Japan. 
Furthermore, the collections are not evenly collected throughout the country, but rather some areas have many more specimens than others (`r s_figure("abundance")`). 
Indeed, one promising avenue of research may be to take advantage of the most densely sampled areas to investigate changes in abundance over time.

@@

@@

# Define nicely formatted terms for printing variable names
resp_var_lookup <- tribble(
  ~resp_var, ~resp_var_pretty,
  "fd_obs_z", "SES of FD",
  "pd_obs_z", "SES of PD",
  "pe_obs_p_upper", "PE *p*-value",
  "rfd_obs_z", "SES of RFD",
  "richness", "Richness",
  "rpd_obs_z", "SES of RPD",
)

@@

  ggplot(plot_data, aes(x = t_value, y = term_print, fill = signif)) +
    geom_vline(xintercept = 0, linetype = "dotted") +
    geom_col() +
    # scale_x_continuous(n.breaks = 4) +
    scale_x_continuous() +
    scale_fill_manual(
      values = c("TRUE" = "grey20", "FALSE" = "grey70"),
      labels = c("TRUE" = "*p* < 0.05", "FALSE" = "*p* > 0.05")) +
    labs(
      subtitle = glue("{unique(plot_data$resp_var_print)},\n{unique(plot_data$model_type_print)}"),
      fill = "Significance") +
    theme_bw(base_size = 9) +
    theme(
      axis.title = element_blank(),
      axis.text.y = element_markdown(),
      legend.text = element_markdown(),
      panel.grid.minor = element_blank(),
      legend.position = "none")
}

@@

**`r figure("coeff")`**. Regression coefficient *t*-values (coefficient divided by standard error) of general mixed models. (a)--(e), environmental models (models including temperature); (f) and (g), reproductive models (models including % apomictic taxa). Statistical significance assessed with a likelihood ratio test (LRT) between the full model and a model with the focal independent variable removed (null model); *p*-value indicates probability of no difference between the full and null model. Name of response variable and model type indicated above each subplot. Independent variable with greatest absolute *t*-value in bold for each model.

@@

e <- ggplot(data, aes(x = taxonomic_cluster, y = pe_obs_p_upper)) +
  geom_jitter(aes(color = taxonomic_cluster)) +
  scale_color_manual(values = bioregion_cols, breaks = 1:4) +
  geom_boxplot(color = "grey80", fill = "transparent", outlier.shape = NA, alpha = 0.25) +
  labs(
    y = "PE *p*-score",
    x = "Cluster",
    color = "Bioregion"
  ) +
  theme_bw() +
  theme(
    axis.title.y = element_markdown(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.y = element_blank(),
  )

@@

**`r figure("div-by-tax-region")`**. Phylogenetic and morphological diversity of the ferns of Japan by taxonomic bioregion. (a), Standard effect size (SES) of phylogenetic diversity (PD). (b), SES of relative PD (RPD). (c), SES of functional diversity (FD). (d), SES of relative FD (RFD). (e), *p*-score of observed phylogenetic endemism (PE) relative to 999 random communities. Only top four clusters with the most sites shown.

@@

`r pagebreak_doc()`

```{r japan-map-doc, fig.cap = "", fig.width = 5, fig.height = 5, eval = params$doc_type == "doc", results = "asis"}
# Print plot if rendering to doc on initial submission
japan_map_plot
cat("\n\n", figure("japan-map"))
```

`r pagebreak_doc()`

```{r raw-diversity-doc, fig.cap = "", fig.width = 7, fig.height = 4, eval = params$doc_type == "doc", results = "asis"}
raw_diversity_plot
cat("\n\n", figure("raw-diversity"))
```

`r pagebreak_doc()`

```{r rand-diversity-doc, fig.cap = "", fig.width = 7, fig.height = 9, eval = params$doc_type == "doc", results = "asis"}
rand_diversity_plot
cat("\n\n", figure("rand-diversity"))
```

`r pagebreak_doc()`

```{r coeff-doc, fig.cap = "", fig.width = 7, fig.height = 5, eval = params$doc_type == "doc", results = "asis"}
coeff_plot
cat("\n\n", figure("coeff"))
```

`r pagebreak_doc()`

```{r modelfit-doc, fig.cap = "", fig.width = 7, fig.height = 5, eval = params$doc_type == "doc", results = "asis"}
modelfit_plot
cat("\n\n", figure("modelfit"))
```

`r pagebreak_doc()`

```{r endemism-doc, fig.cap = "", fig.width = 5, fig.height = 5, eval = params$doc_type == "doc", results = "asis"}
endemism_plot
cat("\n\n", figure("endemism"))
```

`r pagebreak_doc()`

```{r bioregions-doc, fig.cap = "", fig.width = 7, fig.height = 5, eval = params$doc_type == "doc", results = "asis"}
bioregions_plot
cat("\n\n", figure("bioregions"))
```

`r pagebreak_doc()`

```{r div-by-tax-region-doc, fig.cap = "", fig.width = 7, fig.height = 7, eval = params$doc_type == "doc", results = "asis"}
div_by_tax_region_plot
cat("\n\n", figure("div-by-tax-region"))
```

`r pagebreak_doc()`

```{r conserv-status-doc, fig.cap = "", fig.width = 7, fig.height = 4, eval = params$doc_type == "doc", results = "asis"}
conserv_status_plot
cat("\n\n", figure("conserv-status"))
```

`r pagebreak_doc()`

## Data availability statement

Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/japan_ferns_spatial_phy. 
A Docker image to run the code is available at https://hub.docker.com/r/joelnitta/japan_ferns_spatial_phy.
Data files needed to run the analysis and selected results files are available from Dryad at http://doi.org/FIXME:ADD_DOI_HERE_WHEN_ASSIGNED.

## References
<!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->
<div id="refs"></div>

## Biosketch

Joel H. Nitta's research focuses on patterns and processes of diversification in ferns. For more information, see https://joelnitta.com.

Author contributions: J.H.N., B.M., and A.E. conceived the ideas; A.E. provided the data; J.H.N. analyzed the data; W.I provided resources; J.H.N. wrote the manuscript with input from all co-authors.

>>>>

---
title: "Spatial phylogenetics of Japanese ferns: Patterns, processes, and implications for conservation"
mainfont: Roboto-Regular.ttf
fontsize: 11pt
mainfontoptions: 
  - BoldFont=Roboto-Bold.ttf
  - ItalicFont=Roboto-Italic.ttf
  - BoldItalicFont=Roboto-BoldItalic.ttf
output: 
  bookdown::word_document2:
    pandoc_args: [ "--csl", "apa-6th-edition.csl"]
    toc: no
    number_sections: no
    reference_docx: ajb_template.docx
  bookdown::pdf_document2:
    citation_package: default
    toc: no
    number_sections: no
    fig_caption: no
    latex_engine: xelatex
    pandoc_args: [ "--csl", "apa-6th-edition.csl"]
params:
  doc_type: doc
editor_options: 
  chunk_output_type: console
---

@@

Joel H. Nitta^1,5^, Brent D. Mishler^2^, Wataru Iwasaki^1,3^, and Atsushi Ebihara^4^

@@

^5^Author for correspondence: joelnitta@gmail.com

Manuscript received _______; revision accepted _______.

@@

**Premise:** Biodiversity is often measured with a single metric, species richness. However, this ignores evolutionary history and is not sufficient for making conservation decisions. Here, we use a densely sampled dataset to characterize multiple facets and drivers of biodiversity and understand how these relate to bioregions and conservation status.

@@

**Conclusions:** The integrative approach used here reveals previously undetected patterns and drivers of biodiversity in the ferns of Japan. 
The different distributions of each diversity type reflect the major bioregions. 
Future conservation efforts should recognize that threats can vary by biodiversity metric and consider multiple multiple metrics when establishing conservation priorities.

@@

@@

## MATERIALS AND METHODS

@@

## RESULTS

@@

Taxonomic bioregions 3 and 4 have higher SES values of PD, RPD, FD, and RFD than bioregions 1 and 2 (`r figure("div-by-tax-region")`). 
Bioregions 3 and 4 also have consistently high PE *P*-scores, whereas bioregions 1 and 2 have a much wider variation. 
Bioregion 2 has lower RPD and RFD than bioregion 1. 
Comparison of biodiversity metrics across phylogenetic bioregions was similar, except that phylogenetic bioregion 3 mostly corresponds to taxonomic bioregions 3 and 4 combined.

@@

## DISCUSSION

@@

### CONCLUSIONS

@@

Our study integrates diverse, thoroughly collected datasets to characterize the spatial biodiversity of Japanese ferns and gain insight into the processes generating these patterns. We end by summarizing some caveats that must be kept in mind when interpreting our results, while also suggesting avenues for future research.

First, the data include specimens collected over a wide time period, so they do not necessarily represent the current distribution of ferns in Japan. 
Furthermore, the collections are not evenly collected throughout the country, but rather some areas have many more specimens than others (`r s_figure("abundance")`). 
Indeed, one promising avenue of research may be to take advantage of the most densely sampled areas to investigate changes in abundance over time.

@@

## Author contributions

J.H.N., B.M., and A.E. conceived the ideas; A.E. provided the data; J.H.N. analyzed the data; W.I provided resources; J.H.N. wrote the manuscript with input from all co-authors.

## Data availability

Code to replicate all analyses, figures, and this manuscript are available at https://github.com/joelnitta/japan_ferns_spatial_phy. 
A Docker image to run the code is available at https://hub.docker.com/r/joelnitta/japan_ferns_spatial_phy.
Data files needed to run the analysis and selected results files are available from Figshare at http://doi.org/FIXME:ADD_DOI_HERE_WHEN_ASSIGNED.

## Literature cited
<!-- Insert <div id="refs"></div> to place the refs here instead of at the end -->
<div id="refs"></div>

@@

# Define nicely formatted terms for printing variable names
resp_var_lookup <- tribble(
  ~resp_var, ~resp_var_pretty,
  "fd_obs_z", "SES of FD",
  "pd_obs_z", "SES of PD",
  "pe_obs_p_upper", "PE *P*-value",
  "rfd_obs_z", "SES of RFD",
  "richness", "Richness",
  "rpd_obs_z", "SES of RPD",
)

@@

  ggplot(plot_data, aes(x = t_value, y = term_print, fill = signif)) +
    geom_vline(xintercept = 0, linetype = "dotted") +
    geom_col() +
    # scale_x_continuous(n.breaks = 4) +
    scale_x_continuous() +
    scale_fill_manual(
      values = c("TRUE" = "grey20", "FALSE" = "grey70"),
      labels = c("TRUE" = "*P* < 0.05", "FALSE" = "*P* > 0.05")) +
    labs(
      subtitle = glue("{unique(plot_data$resp_var_print)},\n{unique(plot_data$model_type_print)}"),
      fill = "Significance") +
    theme_bw(base_size = 9) +
    theme(
      axis.title = element_blank(),
      axis.text.y = element_markdown(),
      legend.text = element_markdown(),
      panel.grid.minor = element_blank(),
      legend.position = "none")
}

@@

**`r figure("coeff")`**. Regression coefficient *t*-values (coefficient divided by standard error) of general mixed models. (a)--(e), environmental models (models including temperature); (f) and (g), reproductive models (models including % apomictic taxa). Statistical significance assessed with a likelihood ratio test (LRT) between the full model and a model with the focal independent variable removed (null model); *P*-value indicates probability of no difference between the full and null model. Name of response variable and model type indicated above each subplot. Independent variable with greatest absolute *t*-value in bold for each model.

@@

e <- ggplot(data, aes(x = taxonomic_cluster, y = pe_obs_p_upper)) +
  geom_jitter(aes(color = taxonomic_cluster)) +
  scale_color_manual(values = bioregion_cols, breaks = 1:4) +
  geom_boxplot(color = "grey80", fill = "transparent", outlier.shape = NA, alpha = 0.25) +
  labs(
    y = "PE *P*-score",
    x = "Cluster",
    color = "Bioregion"
  ) +
  theme_bw() +
  theme(
    axis.title.y = element_markdown(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.y = element_blank(),
  )

@@

**`r figure("div-by-tax-region")`**. Phylogenetic and morphological diversity of the ferns of Japan by taxonomic bioregion. (a), Standard effect size (SES) of phylogenetic diversity (PD). (b), SES of relative PD (RPD). (c), SES of functional diversity (FD). (d), SES of relative FD (RFD). (e), *P*-score of observed phylogenetic endemism (PE) relative to 999 random communities. Only top four clusters with the most sites shown.

@@
